Run-Time Monitoring with Adjustable Overheads Using Dataflow-Guided Filtering

### 2014-09-12 policies.tex
Various edits based on Ed's comments.

In order to use partial monitoring to enable adjustable overheads, we must also
specify a policy for when and which monitoring events are dropped. In this
section, we discuss some of the options and trade-offs for dropping policies.
We split this decision into two components:
\begin{enumerate}
  \item When do we need to drop events in order to enable reduced overheads?
  \item Which events should be valid for dropping?
\end{enumerate}

Section~\ref{sec:policies.when} discusses options for deciding when events are
dropped while Section~\ref{sec:policies.which} discusses trade-offs for which
events are capable of being dropped.

@@

In this section, we discuss two possible ways to determine when events should be dropped.
The first possibility is to probabilistically drop events. By setting the probability of
dropping events appropriately, overheads can be reduced. This is works well for
enabling partial monitoring for cooperative testing and debugging since the
randomness allows different users and runs to monitor different portions of the
program. However, using a probabilistic dropping policy can make it difficult
to meet a target overhead without prior profiling.

@@

% Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
% keeping track of slack. 
Slack can be easily measured in hardware by using a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, the measured slack
is decremented. This can be difficult to determine precisely due to the
difficulty in measuring certain overheads such as contention for shared memory.
However, we have found that using only the stalls due to FIFO back pressure
works well in practice.

@@

In addition to deciding when dropping is required, trade-offs also exist in
deciding which dropping events can be dropped.
The simplest policy is to drop all monitoring events as
long as slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:policies.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be filtered. Square
nodes indicate events where monitoring checks are performed. In the
example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results were not used for any monitoring checks.

@@

An alternative dropping policy which eliminates this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-only dropping} and we refer to the previous policy of 
dropping any event as \emph{unrestricted dropping}.
Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. We expect source dropping 
to work well when there are a large number of independent metadata flows.
In addition, using a probabilistic dropping policy works poorly with
unrestricted dropping since each event in a chain needs to be randomly selected
for monitoring in order for the check to be performed. Instead, probabilistic
dropping works better when coupled with source-only dropping.

>>>>

In order to use partial monitoring to enable adjustable overheads, we must also
specify a policy for when and which monitoring events are dropped. In this
section, we discuss some of the options and trade-offs for dropping policies.
We split this decision into two components:
\begin{enumerate}
  \item When do we need to drop events in order to enable reduced overheads?
  \item Which events should be dropped?
\end{enumerate}

Section~\ref{sec:policies.when} discusses options for deciding when events are
dropped while Section~\ref{sec:policies.which} discusses trade-offs for which
events can be dropped.

@@

In this section, we discuss two possible ways to determine when events should be dropped.
The first possibility is to probabilistically drop events. By setting the probability of
dropping events appropriately, overheads can be reduced. This works well for
enabling partial monitoring for cooperative testing and debugging since the
randomness allows different users and runs to monitor different portions of the
program. However, using a probabilistic dropping policy can make it difficult
to meet a target overhead without prior profiling.

@@

% Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
% keeping track of slack. 
Slack can be easily measured in hardware by using a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, the measured slack
is decremented. It is difficult to perfectly determine when the main core is impacted by monitoring due to the
difficulty in measuring certain overheads such as contention for shared memory.
However, we have found that using only the stalls due to FIFO back pressure
works well in practice.

@@

In addition to deciding when dropping is required, trade-offs also exist in
deciding which events should be dropped.
The simplest policy is to drop any monitoring events when slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:policies.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be dropped. Square
nodes indicate events where monitoring checks are performed. In the
example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results were not used for any monitoring checks.
That is, by the time we decide to drop event {\tt E}, we have already 
updated metadata due to events {\tt C} and {\tt D} which is no longer needed.

@@

An alternative dropping policy which eliminates this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-only dropping} and we refer to the previous policy of 
dropping any event as \emph{unrestricted dropping}.

Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. We expect source-only dropping 
to work well when there are a large number of independent metadata flows.
In addition, using a probabilistic dropping policy works poorly with
unrestricted dropping. Since each event in a dependent chain (e.g., events {\tt
A} through {\tt E}) needs to be monitored in order for the monitoring check to
be useful, randomly dropping any single event will cause the final check to be
invalid.

For applications where closely matching an overhead target is important, a
slack-based, unrestricted dropping policy is appropriate. However, if matching
the overhead target is not as important, then a slack-based, source-only dropping
policy could provide better coverage. 
Finally, if the goal is to use partial monitoring to enable cooperative
debugging and testing with very low overheads, then a probabilistic,
source-only dropping policy can be used to provide good total coverage over
multiple runs.
